---
title: "Project 4"
output: pdf_document
date: "2023-09-26"
---

```{r setup, include=FALSE, tidy=TRUE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE, 
                      autodep=TRUE, cache.comments=FALSE,
                      message=FALSE, warning=FALSE,
                      fig.width=4.5, fig.height=3)
```

This exercise involves the Boston housing data set in ISLR2. Assume that we are
interested in median home values, medv.
1. Examine medv as a function of crim, zn and indus in a multiple linear regression.
```{r}
library(MASS)
View(Boston)
```

```{r}
lm_all <- lm(medv~crim+zn+indus, Boston)
lm_all
```
A. Identify predictors significant at 5% level.
```{r}
df_all<-df.residual(lm_all)

t_crim<-coef(lm_all)[2]/sqrt(vcov(lm_all)[2,2]) 
t_crim
p_crim<-2*pt(-abs(t_crim),df_all)
p_crim

t_zn<-coef(lm_all)[3]/sqrt(vcov(lm_all)[3,3]) 
t_zn
p_zn<-2*pt(-abs(t_zn),df_all)
p_zn

t_indus<-coef(lm_all)[4]/sqrt(vcov(lm_all)[4,4]) 
t_indus
p_indus<-2*pt(-abs(t_indus),df_all)
p_indus

summary(lm_all)
```
crime, industry, and zoning are all significannt predictors of median home values.

B. List the hypotheses tested in 1.A and their conclusions.

$H_0: \beta_1=0$ false
$H_0: \beta_2=0$ false
$H_0: \beta_3=0$ false

C. What do the estimated regression coefficients in 1.A mean in lay terms? Do
they make sense?

For a neighborhood with no industry and no crime, we can expect 
an increase of .05850 in median home values for an increase of 1 unit in zoning for lots over 25,000 sq. ft.

For a neighborhood with no industry and no zoning for lots over 25,000 sq. ft, we can expect a decrease of .24863 in median home values for an increase of 1 unit in per capita crime.

For a neighborhood with no crime and no zoning for lots over 25,000 sq. ft, we can expect a decrease of .41558 in median home values for an increase of 1 unit in industry.

D. Construct 95% confidence intervals for βcrim, βzn and βindus. How do the
confidence intervals correspond to 1.A and 1.B?
```{r}
#beta_crim <- lm(medv~crim, Boston)
ncrim <- dim(Boston)[[1]]
lowCIcrim <- -0.24863-t_crim*sqrt(var(Boston$crim)/(ncrim-1))
lowCIcrim
upCIcrim <- -0.24863+t_crim*sqrt(var(Boston$crim)/(ncrim-1))
upCIcrim

nzn <- dim(Boston)[[2]]
lowCIzn <- 0.05850-t_zn*sqrt(var(Boston$zn)/(nzn-1))
lowCIzn
upCIzn <- 0.05850+t_zn*sqrt(var(Boston$zn)/(nzn-1))
upCIzn

#nindus <- dim(Boston)[[3]]
#lowCIindus <- -0.41558-t_indus*sqrt(var(Boston$indus)/(nindus-1))
#lowCIindus
#upCIindus <- -0.41558+t_indus*sqrt(var(Boston$indus)/(nindus-1))
#upCIindus
```

E. Calculate the R2 and R2 adj by hand and report whether this matches the R2
and R2 adj from the output and what they mean?
```{r}

```
2. Compare the model from #1 and a simple linear regression of medv as a
function of zn. Which would you prefer?
```{r}
lm(medv~zn, Boston)
```
The model from #1 is better because it shows a more complete picture of the factors than influence median home values. When crime and industry are taken into account, zoning has a less significant effect on home values.