---
title: "Project 1"
author: Sol and Keqing
date: "2023-09-05"
output: pdf_document
---

```{r setup, include=FALSE, tidy=TRUE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE, 
                      autodep=TRUE, cache.comments=FALSE,
                      message=FALSE, warning=FALSE,
                      fig.width=4.5, fig.height=3.5)

library(faraway)
data("hprice")
```

-   Note: for each question, be sure to include both R code and output that is pertinent to your answer(s).

-   Use hprice in the faraway package.

-   The data include 324 observations from 36 US metropolitan statistical areas (MSAs) over 9 years from 1986-1994: 36 x 9 = 324

-   Assume that the MSAs in the data are a simple random sample of the population of MSAs in the US. See <https://www2.census.gov/geo/maps/metroarea/us_wall/Mar2020/CBSA_WallMap_Mar2020.pdf> for MSAs

-   Refer to the R manual for faraway (on Canvas - see Module 1) for background information about this dataset as well as variable definitions.

-   The housing sale price is the outcome variable of interest. Because the dataset has a natural log transformed price variable, narsp, we will recode this to create a variable called "homeprice" by transforming narsp back to the dollar unit for an easier interpretation as follows:

```{r}
hprice$homeprice <- exp(hprice$narsp)*1000
```

1.  What are the mean and the variance of homeprice? What do they mean?

```{r}
mean(hprice$homeprice)
var(hprice$homeprice)
```
#the mean is the average housing sale price 
#the variance is the squared standard deviation and shows the spread from the mean

2.  Construct a 95% confidence interval of the average homeprice. What does the confidence interval imply?

```{r}
t.test(hprice$homeprice, conf.level = 0.95)
```
# We can estimate that the true mean homeprice falls between 90062.70 and 98760.14 95% of the time

3.  Estimate the average homeprice by whether the MSA was adjacent to a coastline, noted in the variable ajwtr, and the standard errors.


```{r}
library(dplyr)
std.error <-function(x) sd(x)/sqrt(length(x))
hprice%>%
  group_by(ajwtr)%>%
  summarize(m=mean(homeprice),
          stderror=std.error(homeprice))
```

4.  Test the difference in homeprice between coastline MSAs and non-coastline MSAs. Clearly state the formula for the null hypothesis, the test method, and your rationale for selecting the method. What do you conclude about the hypothesis?

```{r}
#use t test to determine if mean homeprice in coastline and non-coastline MSAs is the same
#null- mean homeprice of coastal cities= mean homeprice of non-coastal cities

#test equal variance
#null- Variance of homeprice in coastal cities= variance of homeprice in non costal cities
var.test(homeprice ~ ajwtr, hprice, alternative = "two.sided")
#reject null

#use t-test
#null- mean homeprice of costal cities= mean homeprice of non-coastal cities
t.test(homeprice ~ ajwtr, hprice, var.equal=FALSE, conf.int = 0.95)
#reject the null

```

5.  Estimate the Pearson correlation coefficient between homeprice and per capita income of the MSA for a given year, noted in the variable ypc.

```{r}
cor(hprice$homeprice, hprice$ypc, method = 'pearson')
```

6.  Test whether the correlation coefficient between homeprice and ypc is 0, or not. Clearly state the null hypothesis being tested and include the formula.

```{r}
#null- correlation coefficient= 0
cor.test(hprice$homeprice, hprice$ypc)
```

7.  Can you say that per capita income has an effect on the home sales price using the results from #6? Why or why not?

```{r}
#yes, there is a positive correlation between per capita income and home sales price
```

8.  Test the normality of homeprice. Would this change your responses to questions 1-7? Why or why not?

```{r}

library(ggplot2)
qqnorm(hprice$homeprice, main="home price", ylab="y_{i:n}", xlab="m_{i:n}") 
qqline(hprice$homeprice, col="red",lwd=2)
```
#homeprice does not follow a normal distribution, therefore answers 1-7 are not accurate. The sample means are innacurate to the population mean