---
title: "Project 10"
output: html_document
date: "2023-11-21"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Use rats data in faraway.

Description: An experiment was conducted as part of an investigation to combat the
effects of certain toxic agents.
Format: A data frame with 48 observations on the following 3 variables:
time: survival time in tens of hours
poison: the poison type - a factor with levels I II III, randomly assigned to
subjects.
treat: the treatment - a factor with levels A B C D, randomly assigned to subjects.
```{r}
library(faraway)
library(ggplot2)
head(rats)
```


Focus on time as the response and treat as the predictor. Plot time as a
function of treat using a jittered scatter plot. Add a mean of time for each level
of treat to the plot. What pattern do you see? Describe this in layman’s terms.
```{r}
#data('rats')

(means <- aggregate(time~treat+poison,rats, mean)) 


ggplot(rats,aes(x=treat, y=time))+
  geom_point(position=position_jitter(width=0.1))+
  stat_summary(fun="mean",colour = "red")+
  geom_text(data = means, aes(label = time, y = time))
```
The third level of poison had the lowest time across each treatment group.
 The treatment group with the lowest mean time was A, followed by C, then D, then B. 

Perform a one-way ANOVA with time as the response and treat as the
predictor. Use aov() as well as lm(). From each, describe your observation about
what can be concluded.
```{r}
summary(aov(time~treat,rats))
anova(lm(time~treat,rats))
summary(lm(time~treat,rats))
```

From the large F value and very small p value in the first two tests, we can reject the null hypothesis, that the mean time across each treatment group is the same, and conclude that there is a significant difference in means across each treatment group. The first two tests show the same thing. 

The third test shows the estimate for each coefficient independently. Based on the t and p values for D and B, we can conclude the means of these groups are significantly different from the mean of the comparison group A. We reject the nnull hypothesis, that there is not a difference in means from the comparison group.
The t and p values for group C do not provide strong evidence to reject the null, so we conclude there is not a significant difference in the mean of group C and the mean of group A.


Perform pairwise comparison using Tukey’s difference test. Which pairs of
treat are different? How does this result compare to the result from #2?
```{r}
TukeyHSD(aov(time~treat,rats))
#t.test(time~treat,subset(rats,treat!="A"))$p.value
#t.test(time~treat,subset(rats,treat!="B"))$p.value
#t.test(time~treat,subset(rats,treat!="C"))$p.value
#t.test(time~treat,subset(rats,treat!="D"))$p.value
```
The pairs that are significantly different are B-A and C-B. We know this because their intervals do not contain zero and their p values are <.05. this is surprising because the test in #2 showed no significant difference between group A and B but did show a difference between group B and C. It is possible the null was falsely rejected in the test in #2 because of the increased chance of type one error introduced by performing multiple comparisons.


How would you describe the structure of this experiment?

This experiment utilized a multi-way treatment structure, testing the effects of treatment level and poison level at the same time. 
This is a balanced experiment, the same number of rats were assigned to each treatment and poison group. There is no control group. 
