---
title: "Project 1"
author: "Your names here"
date: "2023-09-05"
output: pdf_document
---

```{r setup, include=FALSE, tidy=TRUE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE, 
                      autodep=TRUE, cache.comments=FALSE,
                      message=FALSE, warning=FALSE,
                      fig.width=4.5, fig.height=3.5)

library(faraway)
data("hprice")
```

-   Note: for each question, be sure to include both R code and output that is pertinent to your answer(s).

-   Use hprice in the faraway package.

-   The data include 324 observations from 36 US metropolitan statistical areas (MSAs) over 9 years from 1986-1994: 36 x 9 = 324

-   Assume that the MSAs in the data are a simple random sample of the population of MSAs in the US. See <https://www2.census.gov/geo/maps/metroarea/us_wall/Mar2020/CBSA_WallMap_Mar2020.pdf> for MSAs

-   Refer to the R manual for faraway (on Canvas - see Module 1) for background information about this dataset as well as variable definitions.

-   The housing sale price is the outcome variable of interest. Because the dataset has a natural log transformed price variable, narsp, we will recode this to create a variable called "homeprice" by transforming narsp back to the dollar unit for an easier interpretation as follows:

```{r}
hprice$homeprice <- exp(hprice$narsp)*1000
```

1.  What are the mean and the variance of homeprice? What do they mean?

```{r}
mean_homeprice <- mean(hprice$homeprice)
variance_homeprice <- var(hprice$homeprice)
mean_homeprice
variance_homeprice
```

2.  Construct a 95% confidence interval of the average homeprice. What does the confidence interval imply?

```{r}
confidence_interval <- qt(0.975, df=length(hprice$homeprice)-1) * (sd(hprice$homeprice) / sqrt(length(hprice$homeprice)))
lower_bound <- mean_homeprice - confidence_interval
upper_bound <- mean_homeprice + confidence_interval
c(lower_bound, upper_bound)
```

3.  Estimate the average homeprice by whether the MSA was adjacent to a coastline, noted in the variable ajwtr, and the standard errors.

```{r}
coastline_means <- tapply(hprice$homeprice, hprice$ajwtr, mean)
coastline_se <- tapply(hprice$homeprice, hprice$ajwtr, function(x) sd(x) / sqrt(length(x)))
coastline_means
coastline_se
```

4.  Test the difference in homeprice between coastline MSAs and non-coastline MSAs. Clearly state the formula for the null hypothesis, the test method, and your rationale for selecting the method. What do you conclude about the hypothesis?

```{r}
t_test_result <- t.test(homeprice ~ ajwtr, data = hprice)
t_test_result
```

5.  Estimate the Pearson correlation coefficient between homeprice and per capita income of the MSA for a given year, noted in the variable ypc.

```{r}
correlation <- cor(hprice$homeprice, hprice$ypc)
correlation
```

6.  Test whether the correlation coefficient between homeprice and ypc is 0, or not. Clearly state the null hypothesis being tested and include the formula.

```{r}
cor_test_result <- cor.test(hprice$homeprice, hprice$ypc)
cor_test_result
```

7.  Can you say that per capita income has an effect on the home sales price using the results from #6? Why or why not?

```{r}
if (cor_test_result$p.value < 0.05) {
  cat("Yes, per capita income has a significant effect on home sales price.")
} else {
  cat("No, per capita income doesn't have a significant effect on home sales price.")
}
```

8.  Test the normality of homeprice. Would this change your responses to questions 1-7? Why or why not?

```{r}
shapiro_test <- shapiro.test(hprice$homeprice)
shapiro_test
```
