---
title: "Project 2"
author: "Sol Rabine and Keqing Lu"
date: "9/12/2022"
output: pdf_document
number_sections: yes
fontsize: 12pt
---

```{r setup, include=FALSE, tidy=TRUE}
knitr::opts_chunk$set(echo = TRUE,cache=TRUE, 
                      autodep=TRUE, cache.comments=FALSE,
                      message=FALSE, warning=FALSE)
library("dplyr")
#install.packages("tidyr")
library("tidyr")
```

# 1. JWHT Chapter 2. Exercise 5.
What are the advantages and disadvantages of a very flexible (versus a less flexible)
approach for regression or classification? Under what circumstances might a more flexible
approach be preferred to a less flexible approach? When might a less flexible approach be
preferred?

#Very flexible models have the best and most accurate fit to the data itself. They are preffered when there is a large number of observations and when the goal is to make predictions. Very flexible approaches tend to be less legible. A less flexible approach, like a linear approach may be preferred when giving a broad summary of the data, or when you are interested in the average of the data.


#2. Faraway Chapter 2. Exercise 2.
The dataset uswages is drawn as a sample from the Current Population Survey in 1988.
Fit a model with weekly wages as the response and years of education and experience as
predictors in linear regression. Report and give a simple interpretation to the regression
coefficient for years of education. Now fit the same model but with logged weekly wages. Give
an interpretation to the regression coefficient for years of education. Which interpretation is
more natural?
```{r}
library(faraway)
data("uswages")
```

```{r}
pred_wage<-lm(wage~educ+exper,uswages)
summary(pred_wage)
coef(pred_wage)
```

The regression coefficient for education is 51.1753. This means that each increase of one year of education is associated with a 51.1753 unit increase in wage when experience is fixed.

```{r}

log_wage <- log(uswages$wage)
wlogs<- bind_cols(uswages, log_wage)
pred_log_wage<-lm(log_wage~educ+exper,wlogs)
summary(pred_log_wage)
coef(pred_log_wage)
```

increasing education by one changes the log odds by .09506

