---
title: "Project 3"
output: pdf_document
date: "2023-09-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. JWHT Chapter 2. Modified Exercise 10.
This exercise involves the Boston housing data set. Assume that we are interested in per
capita crime rate, crim.

A. Examine crim with summary() and in a histogram.
```{r}
library(MASS)
summary(Boston$crim)
hist(Boston$crim)
```
B. Focus on suburbs with the crime rate above 25. How many suburbs fall into this group?
What are the pupil-teacher ratios like in those suburbs? How about property tax rates? How
about median home values? How do the pupil-teacher ratios, property tax rates and median
home values compare between these suburbs and the remaining suburbs?
```{r}
X <- split(Boston, Boston$crim>25)
lowcrim <- X[[1]]
highcrim <- X[[2]]
```


```{r}
summary(lowcrim$ptratio)
summary(highcrim$ptratio)
```


```{r}
summary(lowcrim$tax)
summary(highcrim$tax)

```


```{r}
summary(lowcrim$medv)
summary(highcrim$medv)
```
There are 495 suburbs below a crime rate of 25, 11 above a crime rate of 25. The higher crime rate suburbs all had the same pupil to teacher ratio, 20.2. The tax rate was also the same across all the high crime suburbs at 666. The home values in these suburbs had a mean of 9.355 and ranged from 5 to 16.3, with more of the values on the lower end of the range. 

For pupil to teacher ratio, compared to the lower crime suburbs, the high crime suburbs were slightly higher on average, but the lower crime suburbs had a greater range. The tax rates were the same across the high crime suburbs and the average was higher than that of the low crime suburbs. The high crime suburbs had a lower median home value on average. The maximum value in the high crime suburbs was very low compared to that of the low crime suburbs.



C. Create a scatter plot of the crime rates and the median home values for 1) all suburbs, 2)
suburbs bounding Charles River, and 3) suburbs not bounding Charles River. What do you
observe?
```{r}
library(ggplot2)
ggplot(Boston, aes(y=crim, x=medv))+
  geom_point()
```
```{r}
R <- split(Boston, Boston$chas == 1)
bounds <- R[[1]]
dnbound <- R[[2]]

ggplot(bounds, aes(y=crim, x=medv))+
  geom_point()
```

```{r}
ggplot(dnbound, aes(y=crim, x=medv))+
  geom_point()+
  xlim(5,50)+ ylim(-5,100)
```
In general, for all the suburbs, lower home values were associated with higher crime rates. There is a slight increase in crime rates at the highest home values. For suburbs that bound the Charles River, there is a larger range of crime rates and the high values are much higher than those of the other group. For suburbs that do not bound the river, crime rates are generally lower, and we see higher crime rates at the lowest and highest home values.

