---
title: "Project5"
output: pdf_document
date: "2023-10-03"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
This exercise involves the Boston housing data set in ISLR2. Assume that we are
interested in median home values, medv.
```{r}
library(MASS)
View(Boston)
```

1. Examine medv as a function of chas in a simple linear regression model with
an intercept. Notice the nature of chas for its use in this and following models.
What hypothesis are you testing with each coefficient in lay terms? Given the
results of the hypothesis testing, what do the coefficients mean?
```{r}
#class(chas$Boston)
lmchas <- lm(medv~chas, Boston)
summary(lmchas)
```
chas is a binary predictor variable that tells us if a suburb bounds the Charles River.

The intercept tells us the average median home value for the suburbs that do not bound the river. From the output we know that it is 22.09 (1000$). With the coefficient we are testing if there is a difference between the average median home values for those that bound the river and those that do not. 
From the output we know that there is a 6.35 (1000$) difference between these averages.

$H_0:\gamma_2=0$
$H_0:\gamma_2 dne 0$

2. Examine medv as a function of chas and indus in a multiple linear regression.
What do the coefficients mean?
```{r}
lmchasin <- lm(medv~indus+chas, Boston)
summary(lmchasin)
```
For neighborhoods that do not bound the Charles River and do not have industry, the average median home value is the intercept, 29.43 (1000$). For those that do not bound the river, average median home value decreases by .67 (1000 dollars) for increases of 1 unit in industry. Holding industry constant, suburbs that do bound the river on average have a 7.48 (1000 dollar) higher median home value.

3. Given results from #1 and #2, what do you conclude about chas and
indus in relation to medv? Support your conclusion with a visualization.
```{r}
int<-ggplot(aes(x=indus,y=medv),data=Boston)+
  facet_grid(~chas)+
  geom_smooth(method="lm", se=FALSE) 
int

```
Both industry and bounding the Charles river have a significant impact on average median home values, but the effect of industry is larger when a suburb does not bound the river. 
