---
title: "Project 6"
output: pdf_document
date: "2023-10-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(ggplot2)
library(faraway)
```

Faraway Chapter 4. Exercise 2. Using the teengamb data, fit a model with gamble
as the response and the other variables as predictors.
```{r}
data("teengamb")
lm(gamble~sex+status+income+verbal, teengamb)
```
A. Predict the amount that a male with average (given these data) status, income
and verbal score would gamble along with an appropriate 95% CI.

```{r}
mod1<-lm(gamble~sex+status+income+verbal, teengamb)
summary(mod1)
mod1_x<-model.matrix(mod1) 
dim(mod1_x)
head(mod1_x)
```

```{r}
datnew <- data.frame(
  sex = 0,
 status = mean(mod1$model$status),
income = mean(mod1$model$income),
verbal = mean(mod1$model$verbal)
)

predictions <- predict(
  mod1,
  newdata = datnew,
  type = 'response',
  se = TRUE
)
print(predictions)

df_preds = data.frame(datnew, predictions) %>%
  mutate(lower = fit - 1.96 * se.fit,
         upper = fit + 1.96 * se.fit)%>%
  summarise(mn_beta = mean(fit),
            lcl = mean(lower),
            ucl = mean(upper))
print(df_preds)

#predict(mod1, new=datnew, interval="prediction")
```


B. Repeat the prediction for a male with maximal values (for this data) of status,
income and verbal score. Which CI is wider and why is this result expected?
```{r}
datnew1 <- data.frame(
  sex = 0,
 status = max(mod1$model$status),
income = max(mod1$model$income),
verbal = max(mod1$model$verbal)
)

predictions1 <- predict(
  mod1,
  newdata = datnew1,
  type = 'response',
  se = TRUE
)
print(predictions1)

df_preds1 = data.frame(datnew1, predictions1) %>%
  mutate(lower = fit - 1.96 * se.fit,
         upper = fit + 1.96 * se.fit)%>%
  summarise(mn_beta = mean(fit),
            lcl = mean(lower),
            ucl = mean(upper))
print(df_preds1)
```
The CI for a male with maximum values is wider. This is because the residuals for data points at the extreme values are 

C. Fit a model with sqrt(gamble) as the response but with the same predictors.
Now predict the response and give a 95% prediction interval for the individual
in #A. Take care to give your answer in the original units of the response.
```{r}
mod2 <- lm(sqrt(gamble)~sex+status+income+verbal, teengamb)

datnew3 <- data.frame(
  sex = 0,
 status = mean(mod2$model$status),
income = mean(mod2$model$income),
verbal = mean(mod2$model$verbal)
)

predictions3 <- predict(
  mod2,
  newdata = datnew3,
  type = 'response',
  se = TRUE
)
print(predictions3)


predict(mod2, new=datnew3, interval="prediction")

```
The amount that a male with average status, income, and verbal score would spend on gambling yearly is between -.245^2 pounds and 8.344^2

the square root of the pounds that a male with average status, income, and verbal score would spend on gambling yearly is between -.245 and 8.344


D. Repeat the prediction for the model in #C for a female with status=20,
income=1, verbal=10. Comment on the credibility of the result.
```{r}
mod2 <- lm(sqrt(gamble)~sex+status+income+verbal, teengamb)

datnew4 <- data.frame(
  sex = 1,
 status = 20,
income = 1,
verbal = 10
)

predictions4 <- predict(
  mod2,
  newdata = datnew4,
  type = 'response',
  se = TRUE
)
print(predictions4)

predict(mod2, new=datnew4, interval="prediction")
```

This result is less credible than that in question C. because the fit is -2.086, indicating a poor fit.



E. Now, go back to #A. Consider males with average (given these data) status,
income and verbal score would gamble along with an appropriate 95% CI. How
does it compare to the interval in #A?
```{r}


```

